The four postulates of quantum mechanics are three

### 2020-11-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Version submitted to PhysLetters

	\begin{remark}
		The above proposition is, for the most part, an adaptation of the fundamental theorem of projective geometry \cite{fun}. The conservation of the probability imposes the semi-linear map to be either linear or anti-linear (i.e. conjugate-linear).
	\end{remark}

@@

	\begin{remark}
		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear)  has a clear physical interpretation. Since, for a Hilbert space, the conjugate vector space is equivalent to the dual space, a conjugate-linear map is one that preserves the inner product but maps ket vectors into bra vectors. Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system.

		Such a conjugate-linear map is physically indistinguishable from a map that maps kets into kets: the physics is only contained in the probabilities which are the square moduli of inner products. The idea is that, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear. With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.

@@

	\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
		Let $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be a map such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$. Then $m$ must be bilinear. That is:
		\begin{align}
		m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b) \\
		m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)
		\end{align}
		for all $a, a_1, a_2 \in \mathcal{A}$, $b, b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
	\end{prop}

@@

	\begin{references}
		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
		181 (2009).
		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
		(Clarendon Press, Oxford, 1966).
		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
		Quantum Mechanics (Princeton Univ.  Press, 1955).
		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
		(Addison-Welsey, 1968), pg.~176.
		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
		p. 91.
		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
		quantum-mechanical description of physical reality be considered
		complete?, Phys. Rev. {\bf 47}, 777 (1935).
		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
			311}, 350 (2004).
		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
		measurement postulates of quantum mechanics are operationally
		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
		\bibitem{wootters}W.K. Wootters, Optimal Information Transfer and
		Real-Vector-Space Quantum Theory. In: Chiribella G., Spekkens R.
		(eds) Quantum Theory: Informational Foundations and Foils,
		Fundamental Theories of Physics, vol 181. Springer, Dordrecht
		(2016).
		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
		and Quantum Information (Cambridge University Press, Cambridge,
		2000).
		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
		development (World Scientific, 2014).
		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
		free orthodistributive product of orthomodular lattices, Acta Sci.
		Math. (Szeged), {\bf 37}, 263 (1975).
		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
		using the tensor product to describe two quantum systems as one
		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
		(2015).% https://doi.org/10.1007/JHEP12(2015)099
		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
		rule, Phys. Rev. {\bf 155}, 1428 (1967).
		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
		Lett {\bf 87}, 077901 (2001).
		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
		{\bf 92},060402 (2004).
		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
		press, 1961).
		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
		Field Theory (Princeton Univ. Press, 1997).  
		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
			21}, 43 (1991).
		\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
		quantum theory, (North Holland, 1982).
	\end{references}

>>>>

	\begin{remark}
		The above proposition is, for the most part, an adaptation of the fundamental theorem of projective geometry \cite{fun}. The conservation of the probability imposes the semi-linear map to be either linear or anti-linear (i.e. conjugate-linear). This is not new, it is essentially Wigner's theorem, but the proof we offer is insightful as it clearly shows the connection between the construction of the map and the choice of gauge.
	\end{remark}

@@

	\begin{remark}
		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear) corresponds, in physics terms, to a choice of convention. As analogies: a change in metric signature in relativity would change the mathematical space but not the physics; in classical phase-space, a change in signature of the symplectic form would change the mathematical space, but not the physics it represents. These choices are widely recognized as a matter of personal preference.

		In simple terms, for a Hilbert space, the conjugate
                vector space is equivalent to the dual space, so we
                could equivalently choose one or the other. For
                example, Schr\"odinger in the papers in which he
                introduces the Schr\"odinger equation, he writes it
                with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
                the Hilbert space, the other to the dual space, namely
                to the complex-conjugate wave function \cite{sch}. So we can think of the anti-linear map as one that preserves the inner product but maps ket vectors into bra vectors. Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system. This would only make the representation of the composite physical system more complicated, as we need to keep track of the different conventions in the different subspaces. Therefore, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear. With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.

Another way to look at this is that Hermitian operators, and therefore all 
 the physics, are invariant under an anti-linear transformation.
 In contrast, anti-Hermitian operators will change sign. This 
 changes the connection between the generators and the generated 
 transformations (i.e.~while $A$ generates $exp(\frac{Ab}{\imath\hbar})$ 
 on one space, the mapped $A$ will generate 
 $exp(-\frac{Ab}{\imath\hbar})$ in the mapped space). Note that the 
 choice of whether to put the minus or not is arbitrary as long as one is consistent across all generators and transformations.
 Similarly, we typically define $[A, B] = AB - BA$ but we could have 
 alternatively chosen $[A, B] = BA - AB$. The anti-linear map is simply 
 a change of that convention.

@@

	\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
		Given $M$ in \ref{def_comp}, if we can find an $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ such that for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$, then $m$ must be bilinear. That is:
		\begin{align}
		m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b) \\
		m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)
		\end{align}
		for all $a, a_1, a_2 \in \mathcal{A}$, $b, b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
	\end{prop}

@@

	\begin{references}
		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
		181 (2009).
		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
		(Clarendon Press, Oxford, 1966).
		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
		Quantum Mechanics (Princeton Univ.  Press, 1955).
		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
		(Addison-Welsey, 1968), pg.~176.
		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
		p. 91.
		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
		quantum-mechanical description of physical reality be considered
		complete?, Phys. Rev. {\bf 47}, 777 (1935).
		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
			311}, 350 (2004).
		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
		measurement postulates of quantum mechanics are operationally
		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
		\bibitem{wootters}W.K. Wootters, Optimal Information Transfer and
		Real-Vector-Space Quantum Theory. In: Chiribella G., Spekkens R.
		(eds) Quantum Theory: Informational Foundations and Foils,
		Fundamental Theories of Physics, vol 181. Springer, Dordrecht
		(2016).
		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
		and Quantum Information (Cambridge University Press, Cambridge,
		2000).
		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
		development (World Scientific, 2014).
		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
		free orthodistributive product of orthomodular lattices, Acta Sci.
		Math. (Szeged), {\bf 37}, 263 (1975).
		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
		using the tensor product to describe two quantum systems as one
		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
		(2015).% https://doi.org/10.1007/JHEP12(2015)099
		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
		rule, Phys. Rev. {\bf 155}, 1428 (1967).
		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
		Lett {\bf 87}, 077901 (2001).
		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
		{\bf 92},060402 (2004).
		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
		press, 1961).
		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
		Field Theory (Princeton Univ. Press, 1997).  
		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
			21}, 43 (1991).
		\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
		quantum theory, (North Holland, 1982).
		\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81 (1926); English translation in E. Schr\"odinger, {\em Collected papers on Wave Mechanics} (Blackie \& Son, London, 1928).
	\end{references}
